---
title: "MLHW7_JF"
date: March 13, 2022
output: word_document
editor_options:
  chunk_output_type: console
---

```{r libraries, include=FALSE}
library(tidyverse)
library(randomForest)
library(gbm)
library(dplyr)
library(readxl)
library(knitr)
library(Amelia)
library(e1071)
library(rpart)
library(caret)
library(rpart.plot)
library(pROC)
```

# Our main focus is to predict readmission for myocardial infarction.

# Data Prep: Stripping the ID variable, checking for missing data, looking at the frequency of the outcome variable (readmission) to check for balance, and partitioning the data into training and testing using a 70/30 split. 
```{r data_prep}
mi = read.csv("/Users/judyfordjuoh/Desktop/Machine Learning/mi.data.csv") %>%
  janitor::clean_names() %>%
  mutate(readmission = recode(readmission,
                            "0" = "No",
                            "1" = "Yes")) 

mi$readmission <- factor(mi$readmission, levels = c("No", "Yes"))

#Stripping off ID Variable
mi <- mi[,2:16]

#Check distributions, missing data etc, omitting the NAs
summary(mi)
missmap(mi, main = "Missing values vs observed")
#Since there is no missing data we won't do na.omit(mi)

summary(mi$readmission) #Notice that the data is unbalanced so we will have to upsize or downsize. For simplicity sake, I will downsize.

#tidyverse way to create data partition
train_indices <- createDataPartition(y = mi$readmission,p = 0.7,list = FALSE)
train_data <- mi[train_indices, ]
test_data <- mi[-train_indices, ]
```

# REGULARIZED REGRESSION: ELASTIC NET
```{r elastic net} 
#REGULARIZED REGRESSION: ELASTIC NET

set.seed(150)

#Creating 10-fold cross-validation and using up-sampling because of imbalance in data
en.model <- train(
  readmission ~., data = train_data, method = "glmnet",
  trControl = trainControl("cv", number = 10, sampling = "up"), preProc = c("center", "scale"), tuneLength = 10)

#Print the values of alpha and lambda that gave best prediction
en.model$bestTune %>% knitr::kable() # 0.3(alpha)| 0.029915(lambda)| 

#Print all of the options examined. This is a logistic regression = we are using the Accuracy. 
en.model$results %>% knitr::kable()

# Model coefficients
coef(en.model$finalModel, en.model$bestTune$lambda) 

#Confusion Matrix
confusionMatrix(en.model) #0. |(accuracy) 
```
The accuracy of the elastic net model was 62%.

# ENSEMBLE: BAGGING
```{r, bagging using caret calling on randomforest package}
set.seed(150)

#Set our value for mtry hyperparameter (the number of features eligible for selection at each node)
#Remember, in bagging, all predictor features are eligible for selection at each node
mtry.val1 <- expand.grid(.mtry = ncol(train_data) - 1)

trControl_bag = trainControl("cv", number = 10, sampling = "up")

bag_readmission <- train(readmission ~., data = train_data, method = "rf", metric = "Accuracy", trControl = trControl_bag, tuneGrid = mtry.val1, ntree = 100)

# accuracy results was a little bit better than the classification tree accuracy we ran before

bag_readmission$results #mtry: 14, Accuracy: 
varImp(bag_readmission) #wbc , age , esr , sodium 
plot(varImp(bag_readmission))
confusionMatrix(bag_readmission) #Accuracy: 

```
The accuracy of the bagging model was 88.25%. The top three most important variables were white blood cell count (100) age at initial MI (77.60) and liver enzymes (59.14). The least important variables were asthma(0.7179), presence of arrythmia(0.00), and obesity (3.48).

# ENSEMBLE: RANDOM FOREST
```{r}
set.seed(150)

#Trying three different values of mtry (square root, half)
#Set our value for mtry hyperparameter (the number of features eligible for selection at each node)
# since we are not specifying our cross validation, the default is a bootstrap. R is bootstrapping 25 times.

mtry.vals <- c(ncol(train_data) - 1, sqrt(ncol(train_data) - 1), 0.5*ncol(train_data) - 1)

mtry.grid <- expand.grid(.mtry = mtry.vals)

trControl_rf = trainControl("cv", number = 10, sampling = "up")

rf_readmission <- train(readmission ~., data = train_data, method = "rf", metric = "Accuracy", trControl = trControl_rf, tuneGrid = mtry.grid, ntree = 100)

confusionMatrix(rf_readmission) #Accuracy (average) : 
rf_readmission$results
rf_readmission$bestTune #mtry = 6.5
rf_readmission$finalModel

varImp(rf_readmission) #age, wbc , esr 
plot(varImp(rf_readmission))

varImpPlot(rf_readmission$finalModel)

```
The accuracy of the random forest model was 89.5%. The top three most important variables were age at initial MI (90.49), white blood cell count (100), and erythrocyte sedimentation rate (71.8). The least important variables were presence of arrythmia (0.00), asthma (1.44), and obesity (3.98).

## Selecting an "optimal" model and calculate final evaluation metrics in the test set.
The optimal model I selected was the random forest model because the accuracy was the highest out of the three models I've constructed.

```{r EN_optimal_model, results='hide'}
#Checking out info about final model
rf_readmission$finalModel
```

```{r EN_optimal_model continued}
#Make predictions in testset
rf_pred_test <- predict(rf_readmission, test_data)

#Get evaluation metrics from test set
confusionMatrix(rf_pred_test, test_data$readmission, positive = "Yes") #Accuracy   #Sensitivity: #Specificity:

#Create ROC Curve for Analysis
pred.prob <- predict(rf_readmission, test_data, type = "prob")

#Another potential evaluation: Area under the Receiver Operating Curve (AUROC)
#The ROC curve shows the trade-off between sensitivity (or TPR) and specificity (1 – FPR). Classifiers that give curves closer to the top-left corner indicate a better performance. The closer the curve comes to the 45-degree diagonal of the ROC space, the less accurate the test.
analysis <- roc(response = test_data$readmission, predictor = pred.prob[,2])
plot(1 - analysis$specificities,analysis$sensitivities,type = "l",
ylab = "Sensitivity",xlab = "1-Specificity",col = "black",lwd = 2,
main = "ROC Curve for Readmission Classification")
abline(a = 0,b = 1)
```
The accuracy of the random forest model on the test data was 89.59 % with a sensitivity of 0.00 and a specificity of 0.98701.
